Spatial Analysis

### 2022-02-04 Scripts/spatial_analisis.Rmd
Fixed Figure 5 bug on names swap

# group by state
state_fit <- historic_spp %>% 
  group_by(state,year,region,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  # View()
  left_join(total_fited,
            by = c("year","region","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  group_by(state,region,spp,spatial) %>% 
  mutate(RMean = zoo::rollmean(x = percentage, 
                               10,
                               fill = NA,
                               na.rm=T)
  ) %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(RMean) # Remove NAs from rollmean
  ) %>%  
  left_join(state_lat) %>% 
  mutate(
    spatial = ifelse(spatial == "fp","Fishing ports",
                     ifelse(spatial == "sw","State waters","Difference")
    )
  )

@@

# Plot me
ssp_area <-
  ggplot(state_fit) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = abrev
      # fill = state
    )
  ) +
  geom_area(data = historic_periods,
            aes(
              x = year,
              y = n,
              color = period
            ),
            # fill = "gray90",
            alpha = 0.3,
            stat = "identity"
  ) +
  ylab("Propotion (10 yrs running average)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual("State\n(Latitudinal order)",
                    values = state_pallet,
                    labels = state_fit %>% arrange(desc(order)) %>%  pull(state) %>% unique()
  ) +
  scale_color_manual("",values = "grey90") +
  facet_grid(spp~spatial) +
  scale_x_continuous(breaks = seq(1979,2015,9),limits = c(1979,2015))+
  xlab("Year") +
  my_ggtheme_p(leg_pos = "right", hjust = 0.9,ax_tx_s = 14,leg_tx_s = 16, leg_tl_s = 18,ax_tl_s = 18,facet_tx_s = 16) +
  theme(legend.key.width = unit(1,"line"),
        legend.title.align=0.5)

ggsave(filename = "area_plot_avg_spp_sptl.jpg",
       plot = ssp_area,
       path = my_path("R","Figures"),
       width = 14,
       height = 8
)

>>>>

# group by state
state_fit <- historic_spp %>% 
  group_by(state,year,region,spp,spatial) %>% 
  summarise(state_value = sum(value,na.rm= T), .groups = "drop") %>% 
  # View()
  left_join(total_fited,
            by = c("year","region","spp","spatial")) %>%
  mutate(percentage = state_value/total_value*100) %>% 
  group_by(state,region,spp,spatial) %>% 
  mutate(RMean = zoo::rollmean(x = percentage, 
                               10,
                               align = "left",
                               fill = NA,
                               na.rm=T)
  ) %>% 
  #Only show results for spring
  filter(str_detect(region,"Spring"),
         !is.na(RMean) # Remove NAs from rollmean
  ) %>%  
  left_join(state_lat) %>% 
  mutate(
    spatial = ifelse(spatial == "fp","Fishing ports",
                     ifelse(spatial == "sw","State waters","Difference")
    )
  )

        data_new <- state_fit

        data_new$group <- factor(data_new$state,                 # Relevel group factor
                                 levels = state_lat$state)

@@

# Plot me
ssp_area <-
  ggplot(data_new) +
  geom_area(
    aes(
      x = year,
      y = round(RMean),
      fill = group
      # fill = state
    )
  ) +
  geom_area(data = historic_periods,
            aes(
              x = year,
              y = n,
              color = period
            ),
            # fill = "gray90",
            alpha = 0.3,
            stat = "identity"
  ) +
  ylab("Propotion (10 yrs running average)") +
  # viridis::scale_fill_viridis(discrete = T, alpha = 0.8) +
  scale_fill_manual("State\n(Latitudinal order)",
                    values = state_pallet#,
                    # labels = state_fit %>% arrange(desc(order)) %>%  pull(state) %>% unique()
  ) +
  scale_color_manual("",values = "grey90") +
  facet_grid(spp~spatial) +
  scale_x_continuous(breaks = seq(1979,2015,9),limits = c(1979,2015))+
  xlab("Year") +
  my_ggtheme_p(leg_pos = "right", hjust = 0.9,ax_tx_s = 14,leg_tx_s = 16, leg_tl_s = 18,ax_tl_s = 18,facet_tx_s = 16) +
  theme(legend.key.width = unit(1,"line"),
        legend.title.align=0.5)

ggsave(filename = "area_plot_avg_spp_sptl.jpg",
       plot = ssp_area,
       # path = my_path("R","Figures"),
       width = 14,
       height = 8
)
